### 2017-06-25 paper/compositional.tex
expanded the section on compositionality

\begin{definition}[PV-supermartingale]
A 1-dimensional propositionally linear map $\lem$ is a PV supermartingale (PVSM) for a program $\program$ if there exists $\eps>0$ such that $\lem$ is $\eps$-ranking an non-negative in each location $\loc\in\slice{(\program)}$ and unaffected in each $\loc\in\loops(\program)$.
\end{definition}

@@

To overcome the rather strict

>>>>

\begin{definition}[PV-supermartingale {\cite[Definition 7.1.]{HolgerPOPL}}]
A 1-dimensional propositionally linear map $\lem$ is a PV supermartingale (PVSM) for a program $\program$ if there exists $\eps>0$ such that $\lem$ is $\eps$-ranking an non-negative in each location $\loc\in\slice{(\program)}$ and unaffected in each $\loc\in\loops(\program)$.
\end{definition}

As a matter of fact, the condition that $\lem$ should be non-negative in locations of $\slice{(\program)}$ is not explicitly mentioned in \cite{HolgerPOPL}. However, it is implicitly used in some of the proofs and one can easily construct an example where, if the non-negativity in $\slice{(\program)}$ is not required, the Theorem~\ref{thm:holger-comp} below, which also comes from~\cite{HolgerPOPL}, does not hold. Hence, we state the condition explicitly.

@@

To overcome the rather strict typechecking, we use the results on LexRSMs to define a new notion of compositional ranking supermartingales, which we call \emph{nob-negative compositional (NC)} supermartingales. For the sake of generality, we allow NC martingales to be general measurable maps, not necessarily propositionally linear.

\begin{definition}
\label{def:nonneg-comp}
A 1-dimensional measurable map $\lem$ is an NC supermartingale (NCSM) for a program $\program$ if there exists $\eps>0$ such that $\lem$ is:
\begin{compactenum}
\item  non-negative in each location of $\pCFG_{\program}$;
\item 
 $\eps$-ranking in each location $\loc\in\slice{(\program)}$; and
\item 
  unaffected in each $\loc\in\loops(\program)$.
\end{compactenum}
A (propositionally) linear NCSM (LinNCSM) is a NCSM which is also a (propositionally) linear expression map.
\end{definition}

We can prove that NCSMs are a sound method for proving a.s. termination in a compositional way, without any additional assumptions.

\begin{theorem}
Let $\program$ be a \PP{} of the form $\textbf{while } \Psi \textbf{ do } 
\programbody \textbf{ od}$. Assume that each nested loop of $\program$ terminates almost surely from each reachable configuration, and that there exists a NCSM for $\program$. Then $\program$ terminates almost surely.
\end{theorem}
\begin{proof}
\textbf{[PETR: TBD]}
\end{proof}

Hence, NCSMs effectively trade the uniform integrability condition for non-negativity over the whole program. We believe that the latter condition is substantially easier to impose and check automatically, especially in the case of affine probabilistic programs and (propositionally) linear NCSMs: there synthesizing NCSM entails synthesizing sufficient program invariants (for which there is a good automated tool support~\cite{xxx}) and encoding the non-negativity condition into a collection of linear constraints (as for general LexRSMs in Section~\ref{xxx}).

%% I SHOULD MAYBE EXPLAIN THAT THE NEED TO PERFORM A STATIC ANALYSIS OF THE WHOLE PROGRAM IS 
%% NOT LIMITING, AS SOMETHING SIMILAR NEEDS TO BE DONE FOR PVSM. BUT MAYBE LEAVE THIS FOR REBUTTAL. :)
